---
title: "homework-week-6"
author: "Natalie Robinson"
date: "October 21, 2017"
output: html_document
---

# Homework 6

**1. Write a simple R function, z.prop.test() that can perform one- or two-sample Z-tests for proportion data**

```{r}
z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two.sided", conf.level = 0.95) {
  z <- z.test(x = p1, y = p2, alternative = "two.sided", mu = 0, sigma.x = NULL, sigma.y = NULL, conf.level = 0.95)
  if (n*p > 5) {
    return(z, p.value, conf.int)
  } else {
    print("Warning: do not assume normal distribution")
    return(z, p.value, conf.int)
  }
}
```

**2. The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both  longevity~brain size and  log(longevity)~log(brain size).**

```{r}
library(curl)
library(ggplot2)
f <- curl("https://github.com/fuzzyatelin/fuzzyatelin.github.io/raw/master/AN597_Fall17/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

**Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).**
```{r}
# standard plot

m1 <- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data = d)
m1

slope <- m1$coefficients[2]
intercept <- m1$coefficients[1]
equation <- paste("y=", slope, "x", "+", intercept)

g <- ggplot(data = d, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_text(aes(x = 300, y = 200, label = equation))
g
```

```{r}
# plot of log values

log_longevity <- log(d$MaxLongevity_m)
log_brain_size <- log(d$Brain_Size_Species_Mean)

new_data <- data.frame(log_longevity, log_brain_size)

m2 <- lm(log_longevity ~ log_brain_size, data = new_data)
m2

slope2 <- m2$coefficients[2]
intercept2 <- m2$coefficients[1]
equation2 <- paste("y=", slope2, "x", "+", intercept2)

g <- ggplot(data = new_data, aes(x = log_brain_size, y = log_longevity))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_text(aes(x = 4, y = 4, label = equation2))
g
```

**Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.**

*for the standard plot: the slope is 1.218
*for the log plot: the slope is 0.2341

*Both slopes are positive, meaning that as brain size increases, longevity also increases. 
*The slopes are not equal to 0, so the null hypothesis should be rejected. We should accept the alternative hypothesis. 

```{r}
# confidence interval for standard plot
ci1 <- confint(m1, level = 0.90)
ci1
```

```{r}
# confidence interval for log plot
ci2 <- confint(m2, level = 0.90)
ci2
```

**Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.**

```{r}
# standard plot
pi <- predict(m1, newdata = data.frame(brain_size = d$Brain_Size_Species_Mean), interval = "prediction", level = 0.90)
head(pi)

df <- data.frame(cbind(d$Brain_Size_Species_Mean, d$MaxLongevity_m, pi))
names(df) <- c("x", "y", "fit", "lwr", "upr")

g <- ggplot(data = df, aes(x = x, y = y)) + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_line(data = df, aes(x = x, y = lwr), colour = "red")
g <- g + geom_line(data = df, aes(x = x, y = upr), colour = "red")
g <- g + scale_colour_manual("",
                             breaks = c("pi", "ci"),
                             values = c("red", "blue")) #can't figure out how to add a legend
g <- g + geom_text(aes(x = 300, y = 200, label = equation), parse = FALSE)
g
```
Legend: blue = CI, red = PI

```{r}
# log plot
pi2 <- predict(m2, newdata = data.frame(brain_size = log(d$Brain_Size_Species_Mean)), interval = "prediction", level = 0.90)
head(pi2)

df2 <- data.frame(cbind(log(d$Brain_Size_Species_Mean), log(d$MaxLongevity_m), pi2))
names(df2) <- c("x", "y", "fit", "lwr", "upr")

g <- ggplot(data = df2, aes(x = x, y = y)) + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_line(data = df2, aes(x = x, y = lwr), colour = "red")
g <- g + geom_line(data = df2, aes(x = x, y = upr), colour = "red")
g <- g + geom_text(aes(x = 4, y = 4, label = equation2), parse = FALSE)
g
```
Legend: blue = CI, red = PI

**Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?**

```{r}
# standard

ci <- predict(m1, data.frame(Brain_Size_Species_Mean = 800), interval = "confidence", level = 0.90)
ci
```

```{r}
# log

ci2 <- predict(m2, data.frame(log_brain_size = 800), interval = "confidence", level = 0.90)
ci2           
```
The method used in lm() defaults to ordinary least squares, which creates an equation to tell us how Y varies in response to changes in X. The regression model predicts longevity from brain size. However, I don't think this is the best possible model. Brain size could vary based on longevity, or they could covary. Using other models such as RMA, SMA, or MA could create better predictions. 

**Looking at your two models, which do you think is better? Why?**

I think the log model is better. Graphically, it is easier to look at. The scale of the axes is smaller, and it is easier to see each data point. However, a downside of taking the log of each value is that it makes less sense logically. 